Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Community Detection and Parameter Estimation

### 2021-07-07 summary.tex
updated document

\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{float}
\usepackage{mathtools}
\usepackage{natbib}
\usepackage{parskip}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\setcitestyle{numbers,square,comma}
\usepackage{verbatim}
\usepackage{amsthm}
\usepackage{comment}
\usepackage[]{natbib}
\bibliographystyle{plainnat}
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

@@

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\newcommand{\diag}{\text{diag}}
\newcommand{\tr}{\text{Tr}}
\newcommand{\blockdiag}{\text{blockdiag}}
\newcommand{\indep}{\stackrel{\text{indep}}{\sim}}
\newcommand{\iid}{\stackrel{\text{iid}}{\sim}}
\newcommand{\Bernoulli}{\text{Bernoulli}}
\newcommand{\Betadist}{\text{Beta}}
\newcommand{\BG}{\text{BernoulliGraph}}
\newcommand{\PABM}{\text{PABM}}
\newcommand{\RDPG}{\text{RDPG}}
\newcommand{\GRDPG}{\text{GRDPG}}
\newcommand{\Multinomial}{\text{Multinomial}}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\theoremstyle{definition}
\newtheorem*{example}{Example}
\newpage
\spacingset{1.5} % DON'T change the spacing!

@@

\begin{theorem}[Connecting the PABM to the GRDPG for $K = 2$]
\label{theorem1}
Let
$$X = \begin{bmatrix}
\lambda^{(11)} & \lambda^{(12)} & 0 & 0 \\
0 & 0 & \lambda^{(21)} & \lambda^{(22)}
\end{bmatrix} \quad \text{and} \quad
U = \begin{bmatrix} 1 & 0 & 0 & 0 \\
0 & 0 & 1 / \sqrt{2} & 1 / \sqrt{2} \\
0 & 0 & 1 / \sqrt{2} & - 1 / \sqrt{2} \\
0 & 1 & 0 & 0 \end{bmatrix},$$
where the $\{\lambda^{(kl)}\}$ are as defined in Definition 1.
Then $A \sim \mathrm{GRDPG}_{3, 1}(X U)$ and $B \sim \mathrm{PABM}(\{(\lambda^{(kl)}\})$ are
identically distributed.
\end{theorem}

@@

\citet*{noroozi2019estimation} used Sparse Subspace Clustering (SSC) for
community detection in the PABM. SSC is performed by solving an
optimization problem for each observed point. Given
\(X \in \mathbb{R}^{n \times d}\) with vectors
\(x_i^\top \in \mathbb{R}^d\) as rows of \(X\), the optimization problem
\(c_i = \arg\min_{c} ||c||_1\) subject to \(x_i = X c\) and
\(c^{(i)} = 0\) is solved for each \(i = 1, ..., n\). The solutions are
collected collected into matrix
\(C = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix}^\top\) to
construct an affinity matrix \(B = |C| + |C^\top|\). If each \(x_i\) lie
perfectly on one of \(K\) subspaces, \(B\) describes an undirected graph
consisting of \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if
\(x_i, x_j\) are in different subspaces. If \(X\) instead represents
points near \(K\) subspaces with some noise, a final graph partitioning
step may be performed (e.g., edge thresholding or spectral clustering).

@@

@@

\textbf{Proof of Theorem \ref{theorem1}}. This is given by
straightforward matrix multiplication. It suffices to show that

\[X U I_{3, 1} U^\top X^\top =
\begin{bmatrix}
  \lambda^{(11)} (\lambda^{(11)})^\top & \lambda^{(12)} (\lambda^{(21)})^\top \\
  \lambda^{(21)} (\lambda^{(12)})^\top & \lambda^{(22)} (\lambda^{(22)})^\top
\end{bmatrix}\]

\begin{lemma}
\label{lemma1}
Let $P = V D V^\top$ be the spectral decomposition of the edge probability 
matrix for a PABM. Then $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is 
defined as in Eq.\eqref{eq:xy}.
\end{lemma}

\emph{Proof}. By Theorem 2, \(P = X U I_{p, q} U^\top X^\top\), where
\(X\) is defined as in Eq.~\eqref{eq:xy}), $p = K(K+1)/2$, and $q = K(K-1)/2$. % are given by $p =
% K(K+1)/2$ and $q = are defined as
% in equations (\ref{eq:p}) and (\ref{eq:q}). 
Alternatively we have the eigendecomposition \(P = V D V^\top = V
|D|^{1/2} I_{p, q} |D|^{1/2} V^\top\) and \(|\cdot|^{1/2}\) is applied entry-wise. Thus for some
\(Q \in \mathbb{O}(p, q)\),
\[X U Q = V |D|^{1/2}.\]
Let $R = U Q |D|^{-1/2}$. Then $R$ is {\em invertible} and hence
$$X (X^{\top} X)^{-1} X^{\top} = X R (R^{\top} X^{\top} X R)^{-1}
R^{\top} X^{\top}.$$ 
Furthermore $V = XR$ and since $V$ has orthonormal columns, $V^{\top}
V = I$. We thus conclude
$$X (X^{\top} X)^{-1} X^{\top} = V (V^{\top} V)^{-1} V^{\top} = V V^{\top}$$
as desired.
% Therefore, using the fact that \(U U^\top = I\) and \(V^\top V = I\),
% we have
% \[(V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top = 
% (X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top.\]
% The right-hand side becomes
% \[\begin{split}
%   (X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top & 
%     = X U Q Q^{-1} U^\top (X^\top X)^{-1} 
%     U (Q^\top)^{-1} Q^\top U^\top X^\top \\
%   & = X U U^\top (X^\top X)^{-1} U U^\top X^\top \\
%   & = X (X^\top X)^{-1} X^\top.
% \end{split}\]
% The left-hand side becomes:
% \[\begin{split}
%   (V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top & =
%     V |D|^{1/2} |D|^{-1/2} (V^\top V)^{-1} |D|^{-1/2} |D|^{1/2} V^\top \\
%   & = V V^\top.
% \end{split}\]

\textbf{Proof of Theorem \ref{theorem3}}. By Lemma \ref{lemma1},
\(V V^\top = X (X^\top X)^{-1} X^\top\) where \(X\) is defined as in
(\ref{eq:xy}). Since \(X\) is block diagonal with each block
corresponding to one community, \(X (X^\top X)^{-1} X^\top\) is also a
block diagonal matrix with each block corresponding to a community and
zeros elsewhere. Let $B = n VV^{\top}$. Then $B_{ij} = 0$ whenever
vertices $i$ and $j$ belong to different communities.  
%
%then the \(ij\)\textsuperscript{th} element of
%\(n X (X^\top X)^{-1} X^\top = n V V^\top = B\) is 0.

\textbf{Proof of Theorem \ref{theorem4}}. Let \(V_n\) and \(\hat{V}_n\)
be the $n \times K^2$ matrices whose columns are the eigenvectors of \(P\) and \(A\) corresponding to the
\(K (K + 1) / 2\) most positive and \(K (K - 1) / 2\) most negative
eigenvalues, respectively. From Lemma~5 in
\citeauthor{rubindelanchy2017statistical} we have, for some
\(W \in \mathbb{O}(K^2)\) and \(c > 0\), that
$$\|\hat{V} W - V\|_{2 \to \infty} = O_P \big(\frac{(\log n)^c}{n \sqrt{\rho_n}} \big).$$
We furthermore have \(\|V\|_{2 \to \infty} = O_P(n^{-1/2})\). Let
\((v_n^{(i)})^\top\) and \((\hat{v}_n^{(i)})^\top\) denote the $i$th row
row of \(V_n\) and \(\hat{V}_n\), respectively. Recall
Theorem~\ref{theorem3} and note that $B_{ij} = n(v_n^{(i)})^{\top}
v_n^{(j)}$. Now suppose that vertices $i$ and $j$ belongs to different
communities. Then $B_{ij} = 0$ and we have
% Then for \(i\) and
% \(j\) belonging to different
% communities, we ha using the fact that \((v_n^{(i)})^\top v_n^{(j)} = 0\),
\[\begin{split}
\max_{i, j} |(\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)}| &
= \max_{i, j} |(\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)} -
(v_n^{(i)})^\top v_n^{(j)}| \\
& = \max_{i, j} |(\hat{v}_n^{(i)})^\top W W^\top \hat{v}_n^{(j)} -
(v_n^{(i)})^\top v_n^{(j)}| \\
& \leq \max_{i,j} \Bigl(\|W^{\top} \hat{v}_n^{(i)} - v_n^{(i)}\| \times \|\hat{v}_n^{(j)}\|\bigr)
+ \|W^{\top} \hat{v}_n^{(j)} - v_n^{(j)}\| \times \|v_n^{(i)}\|\Bigr) \\
% & \leq 
% \hat{V}_n W^\top W \hat{V}_n - V_n V_n^\top||_{2 \to \infty} \\
% & = ||2 \hat{V}_n W V_n^\top - 2 V_n V_n^\top + \hat{V}_n W W^\top V_n^\top - 
% 2 \hat{V}_n W V_n^\top + V_n V_n^\top||_{2 \to \infty} \\
% & = ||2 (\hat{V}_n W - V_n) V_n^\top + 
% (\hat{V}_n W - V_n) (\hat{V}_n W - V_n)^\top ||_{2 \to \infty} \\
% & \leq 2 
&\leq \|\hat{V}_n W - V_n\|_{2 \to \infty} \times \|\hat{V}_n\|_{2 \to \infty} +
\|\hat{V}_n W - V_n\|_{2 \to \infty} \times \|V_n\|_{2 \to \infty} \\
& \leq \|\hat{V}_n W - V_n\|_{2 \to \infty}^2 + 2 \|\hat{V}_n W -
V_n\|_{2 \to \infty} \times \|V_n\|_{2 \to \infty}
\\
& = O_P \Big( \frac{(\log n)^c}{n^{3/2} \rho_n^{1/2}} \Big)
\end{split}.\]
Scaling all terms in the above inequality by \(n\), we obtain
$$|n (\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)}| = O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)$$
as desired.

@@

\begin{definition}[Subspace incoherence property \cite{jmlr-v28-wang13}]
\end{definition}

@@

\textbf{Proof of Theorem \ref{theorem5}}. The basis of this proof is
Theorem 6 from \citeauthor{jmlr-v28-wang13}, which states that the
subspace detection property holds if the noise is small enough and the
subspace inradius is greater than the subspace incoherence for each
community \(k\).

@@

Thus the conditions for the subspace detection property from Theorem 6
from \citeauthor{jmlr-v28-wang13} are satisfied with probability 1 as
\(n \to \infty\).

@@

\textbf{Proof of Theorem \ref{theorem6}}. Let \(P\) be
organized by community such that \(P^{(k \ell)}\) denote the $n_k \times
n_{\ell}$ matrix obtained by keeping only the rows of $P$
corresponding to vertices in community $k$ and the columns of $P$
corresponding to vertices in community $\ell$. We define $A^{(k
  \ell)}$ analogously. Recall that $P^{(k \ell)} = \lambda^{(k \ell)} (\lambda^{(\ell k)})^{\top}$ for all $k, \ell$. We now consider estimation of $P^{(k \ell)}$
for the cases when $k = \ell$ versus when $k \not = \ell$.

\emph{Case \(k = l\)}. let $P^{(kk)} =
\sigma_{kk}^2 u^{(kk)} (u^{(kk)})^\top$ be the singular value
decomposition of $P^{(kk)}$. We can then define
$\tilde{\lambda}^{(kk)} = \sigma_{kk} u^{(kk)}$. 
Now let $\hat{U}^{(kk)} \hat{\Sigma}^{(kk)} (\hat{U}^{(kk)})^\top$ be the
singular value decomposition of \(A^{(kk)}\), and let
$\hat{\sigma}_{kk}^2 \hat{u}^{(kk)} (\hat{u}^{(kk)})^\top$ be the
best rank-one approximation of $A^{(kk)}$. Define
\(\hat{\lambda}^{(kk)} = \hat{\sigma}_{kk} \hat{u}^{(kk)}\). Then
\(\hat{\lambda}^{(kk)}\) is the adjacency spectral embedding approximation of \(\lambda^{(kk)}\)
and by Theorem 5 of \citeauthor{rubindelanchy2017statistical}, we have
%adjacency spectral embedding \(\hat{\lambda}^{(kk)}\) approximates
$$\|\hat{\lambda}^{(kk)} - \lambda^{(kk)}\|_{\infty} = \frac{(\log n_k)^c}{\sqrt{n_k}}.$$
Here $\|\cdot\|_{\infty}$ denote the $\ell_{\infty}$ norm of a vector.

@@

where each \(\hat{\lambda}^{(kl)}\) is defined as in Algorithm 3,
approximates the latent positions of \(M^{(kl)}\) up to indefinite
orthogonal transformation by the rate given in Theorem 5 of
\citeauthor{rubindelanchy2017statistical}.\\
In this case, the indefinite orthogonal transformation \(W_*\) in the
GRDPG result \cite{rubindelanchy2017statistical} is of the form
\(U^\top \hat{U}\). The eigenvalues of \(M\) are distinct since the
signature for this GRDPG is \((1, 1)\), and \(U^\top \hat{U}\) is block
diagonal, resulting in \(W_* \stackrel{a.s.}{\to} I\). Therefore, the
adjacency spectral embedding of \(\hat{M}^{(kl)}\) is a direct
estimation of the specific latent positions outlined for \(M^{(kl)}\),
up to sign flip.

\renewcommand\refname{References}
  \bibliography{misc.bib}

>>>>

\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{float}
\usepackage{mathtools}
\usepackage{natbib}
\bibliographystyle{abbrvnat}
\usepackage{parskip}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\setcitestyle{numbers,square,comma}
\usepackage{verbatim}
\usepackage{amsthm}
\usepackage{comment}
\bibliographystyle{plainnat}
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

@@

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{Tr}}
\newcommand{\blockdiag}{\mathrm{blockdiag}}
\newcommand{\indep}{\stackrel{\mathrm{indep}}{\sim}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
\newcommand{\Betadist}{\mathrm{Beta}}
\newcommand{\BG}{\mathrm{BernoulliGraph}}
\newcommand{\PABM}{\mathrm{PABM}}
\newcommand{\RDPG}{\mathrm{RDPG}}
\newcommand{\GRDPG}{\mathrm{GRDPG}}
\newcommand{\Multinomial}{\mathrm{Multinomial}}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\theoremstyle{definition}
\newtheorem*{example}{Example}
\newpage
\spacingset{1.5} % DON'T change the spacing!

@@

\begin{theorem}[Connecting the PABM to the GRDPG for $K = 2$]
\label{theorem1}
Let
$$X = \begin{bmatrix}
\lambda^{(11)} & \lambda^{(12)} & 0 & 0 \\
0 & 0 & \lambda^{(21)} & \lambda^{(22)}
\end{bmatrix} \quad \text{and} \quad
U = \begin{bmatrix} 1 & 0 & 0 & 0 \\
0 & 0 & 1 / \sqrt{2} & 1 / \sqrt{2} \\
0 & 0 & 1 / \sqrt{2} & - 1 / \sqrt{2} \\
0 & 1 & 0 & 0 \end{bmatrix},$$
where the $\{\lambda^{(kl)}\}$ are as defined in Definition 1.
Then $A \sim \mathrm{GRDPG}_{3, 1}(X U)$ and $B \sim \mathrm{PABM}(\{\lambda^{(kl)}\}_K)$ are
identically distributed.
\end{theorem}

\begin{proof}
This is given by
straightforward matrix multiplication.

\[X U I_{3, 1} U^\top X^\top =
\begin{bmatrix}
  \lambda^{(11)} (\lambda^{(11)})^\top & \lambda^{(12)} (\lambda^{(21)})^\top \\
  \lambda^{(21)} (\lambda^{(12)})^\top & \lambda^{(22)} (\lambda^{(22)})^\top
\end{bmatrix}\]
\end{proof}

@@

\citet*{noroozi2019estimation} used Sparse Subspace Clustering (SSC) \cite{5206547} for
community detection in the PABM. SSC is performed by solving an
optimization problem for each observed point. Given
\(X \in \mathbb{R}^{n \times d}\) with vectors
\(x_i^\top \in \mathbb{R}^d\) as rows of \(X\), the optimization problem
\(c_i = \arg\min_{c} ||c||_1\) subject to \(x_i = X c\) and
\(c^{(i)} = 0\) is solved for each \(i = 1, ..., n\). The solutions are
collected collected into matrix
\(C = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix}^\top\) to
construct an affinity matrix \(B = |C| + |C^\top|\). If each \(x_i\) lie
perfectly on one of \(K\) subspaces, \(B\) describes an undirected graph
consisting of \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if
\(x_i, x_j\) are in different subspaces. If \(X\) instead represents
points near \(K\) subspaces with some noise, a final graph partitioning
step may be performed (e.g., edge thresholding or spectral clustering).

@@

\begin{proof}
By Theorem \ref{theorem2}, \(P = X U I_{p, q} U^\top X^\top\), where
\(X\) is defined as in Eq.~\eqref{eq:xy}, $p = K(K+1)/2$, and $q = K(K-1)/2$.
Alternatively we have the eigendecomposition \(P = V D V^\top = V
|D|^{1/2} I_{p, q} |D|^{1/2} V^\top\) and \(|\cdot|^{1/2}\) is 
applied entry-wise. Thus for some \(Q \in \mathbb{O}(p, q)\),

\[X U Q = V |D|^{1/2}.\]

Let $R = U Q |D|^{-1/2}$. Then $R$ is {\em invertible} and hence

$$X (X^{\top} X)^{-1} X^{\top} = X R (R^{\top} X^{\top} X R)^{-1}
R^{\top} X^{\top}.$$

Furthermore $V = XR$ and since $V$ has orthonormal columns, $V^{\top}
V = I$. Thus,

$$X (X^{\top} X)^{-1} X^{\top} = V (V^{\top} V)^{-1} V^{\top} = V V^{\top}$$.

Again by Theorem \ref{theorem2}, $X$ is block diagonal, with each block 
corresponding to a community. Then $n X (X^\top X)^{-1} X^\top$ is also block 
diagonal with each block corresponding to a community, and zeros elsewhere. 
Thus, $[n V V^\top]_{ij} = [n X (X^\top X)^{-1} X^\top]_{ij} = 0$ if vertices 
$i$ and $j$ are from different communities.
\end{proof}

@@

\begin{proof}[Proof of Theorem \ref{theorem4}]
Let \(V_n\) and \(\hat{V}_n\)be the $n \times K^2$ matrices 
whose columns are the eigenvectors of \(P\) and \(A\) corresponding to the
\(K (K + 1) / 2\) most positive and \(K (K - 1) / 2\) most negative
eigenvalues, respectively. From Lemma~5 in
\citeauthor{rubindelanchy2017statistical} we have, for some
\(W \in \mathbb{O}(K^2)\) and \(c > 0\), that
$$\|\hat{V} W - V\|_{2 \to \infty} = O_P \big(\frac{(\log n)^c}{n \sqrt{\rho_n}} \big).$$
We furthermore have \(\|V\|_{2 \to \infty} = O_P(n^{-1/2})\). Let
\((v_n^{(i)})^\top\) and \((\hat{v}_n^{(i)})^\top\) denote the $i$th row
row of \(V_n\) and \(\hat{V}_n\), respectively. Recall
Theorem~\ref{theorem3} and note that $B_{ij} = n(v_n^{(i)})^{\top}
v_n^{(j)}$. Now suppose that vertices $i$ and $j$ belongs to different
communities. Then $B_{ij} = 0$ and we have
% Then for \(i\) and
% \(j\) belonging to different
% communities, we ha using the fact that \((v_n^{(i)})^\top v_n^{(j)} = 0\),
\[\begin{split}
\max_{i, j} |(\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)}| &
= \max_{i, j} |(\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)} -
(v_n^{(i)})^\top v_n^{(j)}| \\
& = \max_{i, j} |(\hat{v}_n^{(i)})^\top W W^\top \hat{v}_n^{(j)} -
(v_n^{(i)})^\top v_n^{(j)}| \\
& \leq \max_{i,j} \Bigl(\|W^{\top} \hat{v}_n^{(i)} - v_n^{(i)}\| \times \|\hat{v}_n^{(j)}\|\bigr)
+ \|W^{\top} \hat{v}_n^{(j)} - v_n^{(j)}\| \times \|v_n^{(i)}\|\Bigr) \\
% & \leq 
% \hat{V}_n W^\top W \hat{V}_n - V_n V_n^\top||_{2 \to \infty} \\
% & = ||2 \hat{V}_n W V_n^\top - 2 V_n V_n^\top + \hat{V}_n W W^\top V_n^\top - 
% 2 \hat{V}_n W V_n^\top + V_n V_n^\top||_{2 \to \infty} \\
% & = ||2 (\hat{V}_n W - V_n) V_n^\top + 
% (\hat{V}_n W - V_n) (\hat{V}_n W - V_n)^\top ||_{2 \to \infty} \\
% & \leq 2 
&\leq \|\hat{V}_n W - V_n\|_{2 \to \infty} \times \|\hat{V}_n\|_{2 \to \infty} +
\|\hat{V}_n W - V_n\|_{2 \to \infty} \times \|V_n\|_{2 \to \infty} \\
& \leq \|\hat{V}_n W - V_n\|_{2 \to \infty}^2 + 2 \|\hat{V}_n W -
V_n\|_{2 \to \infty} \times \|V_n\|_{2 \to \infty}
\\
& = O_P \Big( \frac{(\log n)^c}{n^{3/2} \rho_n^{1/2}} \Big)
\end{split}.\]
Scaling all terms in the above inequality by \(n\), we obtain
$$|n (\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)}| = O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)$$
as desired.
\end{proof}

@@

\begin{definition}[Subspace incoherence property \cite{jmlr-v28-wang13}]
Let $\nu_i(X, \lambda) = 
\arg\max_\eta x_i^\top \eta - \frac{1}{2 \lambda} \eta^\top \eta$ 
subject to $\|X_{-i} \eta\|_\infty \leq 1$. 
Define the projected dual direction of $X$ to subspace $\mathcal{S}$ as 
$v_i(X, \lambda, \mathcal{S}) = 
\frac{\mathbb{P}_{\mathcal{S}}(\nu_i)}{\|\mathbb{P}_{\mathcal{S}}(\nu_i)\|}$.

If $X = \begin{bmatrix} X^{(1)} & \cdots & X^{(K)} \end{bmatrix}^\top$ 
is a set of points lying near $K$ subspaces $\{S^{(1)}, ..., S^{(K)}\}$ 
and each $x^{(k)}_i$ has corresponding point $y^{(k)}_i$ which lies exactly on 
$\mathcal{S}^{(k)}$, let $v_i^{(k)} = v_i(X^{(k)}, \lambda, \mathcal{S}^{(k)})$ 
and 
$V^{(k)} = \begin{bmatrix} v_1^{(k)} & \cdots & v_{n_k}^{(k)} \end{bmatrix}^\top$. 
Then we define $X^{(k)}$ as $\mu_k$-incoherent to $X^{(-k)}$ where

$$\mu_k = \max_{y \in Y^{(-k)}} \|V^{(k)} y\|_\infty$$
\end{definition}

@@

\begin{proof}[Proof of Theorem \ref{theorem5}]
The basis of this proof is
Theorem 6 from \citeauthor{jmlr-v28-wang13}, which states that the
subspace detection property holds if the noise is small enough and the
subspace inradius is greater than the subspace incoherence for each
community \(k\).

@@

Thus the conditions for the subspace detection property from Theorem 6
from \citeauthor{jmlr-v28-wang13} are satisfied with probability 1 as
\(n \to \infty\).
\end{proof}

@@

\begin{proof}[Proof of Theorem \ref{theorem6}]
Let \(P\) be organized by community such that \(P^{(k \ell)}\) 
denote the $n_k \times n_{\ell}$ matrix obtained by keeping only the rows of $P$
corresponding to vertices in community $k$ and the columns of $P$
corresponding to vertices in community $\ell$. We define 
$A^{(k  \ell)}$ analogously. 
Recall that $P^{(k \ell)} = \lambda^{(k \ell)} (\lambda^{(\ell k)})^{\top}$ 
for all $k, \ell$. We now consider estimation of $P^{(k \ell)}$
for the cases when $k = \ell$ versus when $k \not = \ell$.

\emph{Case \(k = l\)}. let $P^{(kk)} =
\sigma_{kk}^2 u^{(kk)} (u^{(kk)})^\top$ be the singular value
decomposition of $P^{(kk)}$. We can then define
$\tilde{\lambda}^{(kk)} = \sigma_{kk} u^{(kk)}$. 
Now let $\hat{U}^{(kk)} \hat{\Sigma}^{(kk)} (\hat{U}^{(kk)})^\top$ be the
singular value decomposition of \(A^{(kk)}\), and let
$\hat{\sigma}_{kk}^2 \hat{u}^{(kk)} (\hat{u}^{(kk)})^\top$ be the
best rank-one approximation of $A^{(kk)}$. Define
\(\hat{\lambda}^{(kk)} = \hat{\sigma}_{kk} \hat{u}^{(kk)}\). Then
\(\hat{\lambda}^{(kk)}\) is the adjacency spectral embedding approximation of \(\lambda^{(kk)}\)
and by Theorem 5 of \citeauthor{rubindelanchy2017statistical}, we have

$$\|\hat{\lambda}^{(kk)} - \lambda^{(kk)}\|_{\infty} = O_P \bigg( \frac{(\log n_k)^c}{\sqrt{n_k}}\bigg).$$

Here $\|\cdot\|_{\infty}$ denote the $\ell_{\infty}$ norm of a vector.

@@

where each \(\hat{\lambda}^{(kl)}\) is defined as in Algorithm 3,
approximates the latent positions of \(M^{(kl)}\) up to indefinite
orthogonal transformation by the rate given in Theorem 5 of
\citeauthor{rubindelanchy2017statistical}.\\
In this case, the indefinite orthogonal transformation \(W_*\) in the
GRDPG result \cite{rubindelanchy2017statistical} is of the form
\(U^\top \hat{U}\). The eigenvalues of \(M\) are distinct since the
signature for this GRDPG is \((1, 1)\), and \(U^\top \hat{U}\) is block
diagonal, resulting in \(W_* \stackrel{a.s.}{\to} I\). Therefore, the
adjacency spectral embedding of \(\hat{M}^{(kl)}\) is a direct
estimation of the specific latent positions outlined for \(M^{(kl)}\),
up to sign flip.
\end{proof}

\renewcommand\refname{References}

\bibliography{misc.bib}
